---
title: "Figma API connection and file extraction"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r development, include=FALSE}
library(testthat)
library(purrr)
library(httr)
library(glue)
library(stringr)
pkgload::load_all()
```

<!--
 You need to run the 'description' chunk in the '0-dev_history.Rmd' file before continuing your code there.
-->

# Connection to Figma API and extraction of the content of a file: `get_file_content()` function

## Connection to Figma API

- Create a personal access token on your Figma account (https://www.figma.com/developers/api Section Access tokens)
- Save this token in .Renviron `usethis::edit_r_environ()`, call the variable `FIGMA_TOKEN`
- Restart R session `rstudioapi::restartSession()`
- Check if the personal access token is registered `Sys.getenv("FIGMA_TOKEN")` 

## Extraction of the content of a file

- Find the key of your Figma file. It corresponds to the character **wRqIvMmymzSPuj0sEhnORb** in the url "https://www.figma.com/file/**wRqIvMmymzSPuj0sEhnORb**/..." (example).

```{r function-get_file_content}
#' Connect to Figma API and extract content of a file.
#'
#' @param file_key Character. Key of the Figma file. 
#' @param acess_token Character. Figma access token.
#'
#' @importFrom httr VERB add_headers content
#' @importFrom stringr str_remove_all
#' @importFrom glue glue
#' @importFrom cli cli_alert_danger cli_alert_success
#'
#' @details
#' The key of your Figma file corresponds to the character 
#' wRqIvMmymzSPuj0sEhnORb in the url 
#' "https://www.figma.com/file/wRqIvMmymzSPuj0sEhnORb/..." (example).
#' The acess token can be created in your Figma account 
#' (https://www.figma.com/developers/api Section Access tokens)
#'
#' @return A list. Content or the file.
#' @export
get_file_content <- function(file_key, acess_token = Sys.getenv("FIGMA_TOKEN")) {
  
  api_response <- VERB(verb = "GET", 
                       url = glue("https://api.figma.com/v1/files/{file_key}"), 
                       add_headers(`X-FIGMA-TOKEN` = acess_token), 
                       encode = "json")
  
  if (api_response$status_code == "404") {
    
    cli_alert_danger("Content of the file has not been extracted. Please check the file key or your access token.")
    stop()
    
  } else {
    
    file_content <- content(api_response)
    
    file_name <- file_content$name %>% 
      str_remove_all("[:punct:]")
    
    cli_alert_success(glue("Content of the file {file_name} has been successfully extracted."))
    
    return(file_content)
    
  }
  
}
```

```{r examples-get_file_content}
file_content <- get_file_content(file_key = "wRqIvMmymzSPuj0sEhnORb",
                                 acess_token = Sys.getenv("FIGMA_TOKEN"))
```

```{r tests-get_file_content}
test_that("get_file_content() works", {

  # No error
  expect_error(file_content <- get_file_content(file_key = "wRqIvMmymzSPuj0sEhnORb",
                                                acess_token = Sys.getenv("FIGMA_TOKEN")), 
               regexp = NA)
  
  expect_error(file_content <- get_file_content(file_key = "wRqIvMmymzSPuj0sEhnORb"), 
               regexp = NA)
  
  expect_equal(length(file_content), 
               12)
  
  # Error
  expect_error(file_content <- get_file_content(acess_token = ""))
  
  expect_error(file_content <- get_file_content(file_key = ""))
  
  expect_error(file_content <- get_file_content(file_key = "", 
                                                acess_token = ""))
  
})
```


```{r development-inflate, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly
fusen::inflate(flat_file = "dev/flat_figma_api_connection.Rmd", 
               vignette_name = "b - Figma API connection and file extraction", 
               open_vignette = FALSE,
               check = FALSE,
               overwrite = TRUE)
```
